---
title: "R Coding Exercise"
output:
  html_document:
    toc: FALSE
---

Placeholder file for the future R coding exercise.

```{r}
#load dslabs package
library("dslabs")
library('tidyverse')
#look at help file for gapminder data
help(gapminder)
#get an overview of data structure
str(gapminder)
#get a summary of data
summary(gapminder)
#determine the type of object gapminder is
class(gapminder)
#select only africa
africadata <- gapminder[which(gapminder$continent=='Africa'),]
#get an overview of data structure
str(africadata)
#get a summary of data
summary(africadata)
#infant_mortality and life_expectancy
moex <- africadata %>% select(c('infant_mortality','life_expectancy'))
#population and life_expectancy
poex <- africadata %>% select(c('population','life_expectancy'))
#get an overview of data structure
str(moex)
str(poex)
#get a summary of data
summary(moex)
summary(poex)

#scatter plot infant mortality VS life expectancy
plot(moex$infant_mortality,moex$life_expectancy, main="infant mortality VS life expectancy", xlab="infant mortality", ylab="life expectancy", pch=19)
```

```{r}
#scatter plot population size VS life expectancy
plot(poex$population,poex$life_expectancy, main="population size VS life expectancy", xlab="population size", ylab="life expectancy", pch=19,log='x')
```

```{r}
#the year which infant_mortality is NA
namor <-africadata$year[which(is.na(africadata$infant_mortality))]
table(namor)
```

```{r}
#2000 year
afri2000 <- africadata[which(africadata$year==2000),]
str(afri2000)
summary(afri2000)

#plot 
plot(afri2000$infant_mortality,afri2000$life_expectancy, main="infant mortality VS life expectancy", xlab="infant mortality", ylab="life expectancy", pch=19)
#plot
plot(afri2000$population,afri2000$life_expectancy, main="population size VS life expectancy", xlab="population size", ylab="life expectancy", pch=19,log='x')
```

```{r}
fit1 <- lm(life_expectancy~infant_mortality,data=afri2000)
summary(fit1)
fit2 <- lm(life_expectancy~population,data=afri2000)
summary(fit2)
```

Here we know infant mortality has statistically significant negative relationship with life-expectancy. But we don't have enough statistical evidence about the relationship between population and life expectancy.

**--------------------------------------------------------------------------------------------------------------------------**

**THIS SECTION ADDED BY SHIWANI SAPKOTA**

**I am interested on looking at infant mortality, life expectancy, fertility, and population variables for European countries in the year 2010 using gapminder data.**

**DATA PROCESSING**

```{r}
# Creating an object europedata containing only the European countries
europedata <- gapminder %>% filter(continent == "Europe")

# Looking at the structure and summary of object (europedata)
str(europedata)
summary(europedata)
glimpse(europedata)
```

```{r}
# Looking at the years having missing data for infant mortality in europedata
infantmort_yrs_europe <- europedata %>% 
                        filter(is.na(infant_mortality))
table(infantmort_yrs_europe$year)
```

```{r}
# Creating a new object named europedata_2010 by using only year 2010 data from europedata
europedata_2010 <- europedata %>%
                   filter(year == 2010)

# Looking at the structure and summary of object (europedata_2010)
str(europedata_2010)
summary(europedata_2010)
glimpse(europedata_2010)
```

```{r}
# Looking at the top 5 European countries having highest infant mortality for the year 2010
europe_infantmort <- europedata_2010 %>% select (country, infant_mortality) %>%
arrange(desc(`infant_mortality`))
europe2010_infantmort <- head(europe_infantmort, 5)
europe2010_infantmort

# Looking at the top 5 European countries having highest life expectancy for the year 2010
europe_lifeexpect <- europedata_2010 %>% select (country, life_expectancy) %>%
arrange(desc(`life_expectancy`))
europe2010_lifeexpect <- head(europe_lifeexpect, 5)
europe2010_lifeexpect

# Looking at the top 5 European countries having largest population size for the year 2010
europe_pop <- europedata_2010 %>% select (country, population) %>%
arrange(desc(`population`))
europe2010_pop <- head(europe_pop, 5)
europe2010_pop
```

**DATA VISUALIZATION**

```{r}
# Creating the object country to sort the 5 European countries having the highest infant mortality in the year 2010
europe2010_infantmort <- europe2010_infantmort %>%
mutate(country = as_factor(country),
country <- fct_reorder(country, infant_mortality))

# Plotting the horizontal barplot for visualizing top 5 European countries with the highest infant mortality in 2010
ggplot(data = europe2010_infantmort, aes(x = country, y = `infant_mortality`)) +
geom_bar(stat = "identity", width = 0.75, color = "blue", fill = "steelblue") +
coord_flip() +
labs(title = "Top 5 European countries with the highest infant mortality in 2010",
x = "Country", y = "Infant Mortality") +
theme_classic() +
theme(axis.text = element_text(face="bold"), plot.title = element_text(hjust = 0.5, size = 14),
axis.title = element_text(size = 14))
```

```{r}
# Plotting fertility as a function of infant mortality in the year 2010 for European countries
ggplot(data = europedata_2010,
       (aes(x = infant_mortality, y = fertility))) +
  theme_classic() +
  geom_point(size = 1.0) +
  labs(title = "Fertility as a function of infant mortality in the year 2010 for European countries",
       x = "Infant Mortality", y = "Fertility") +
  theme(axis.text = element_text(face = "bold"), plot.title = element_text(hjust = 0.5, size = 14),
        axis.title = element_text(size = 14))
```

```{r}
# Plotting life expectancy as a function of population size in the year 2010 for European countries
ggplot(data = europedata_2010,
       (aes(x = population, y = life_expectancy))) +
  theme_classic() +
  geom_point(size = 1.0) +
  scale_x_continuous(trans = 'log10') +
  labs(title = "Life expectancy as a function of population size in the year 2010 for European countries", 
       x = "Population (In Logscale)", y = "Life Expectancy (In Years)") +
  theme(axis.text = element_text(face = "bold"), plot.title =   element_text(hjust = 0.5, size = 14),
        axis.title = element_text(size = 14))
```

**FITTING MODEL**

```{r}
# Using lm function to fit linear regression model using fertility as the outcome and infant mortality as the predictor for the year 2010
fit3 <- lm(fertility ~ infant_mortality, data = europedata_2010)
# Tabulating the output fit3 from lm using broom package
broom::tidy(fit3)

# Using lm function to fit linear regression model using life expectancy as the outcome and population size as the predictor for the year 2010
fit4 <- lm(life_expectancy ~ population, data = europedata_2010)
# Tabulating the output fit4 from lm using broom package
broom::tidy(fit4)
```

**CONCLUSION**

**Based on the p-values for each fit, we found somewhat statistically significant negative correlation between infant mortality and fertility in the year 2010 for European countries (p-value: 0.0429) while we found no statistically significant correlation between population size and life expectancy in the year 2010 for European countries (p-value: 0.345 ).**
